\label{sec:2bound3}

We now have the necessary framework to demonstrate that a closed, oriented 2--manifold bounds a 3--manifold through an explicit construction.
This is useful as a jumping off point for the case one dimension higher, that of 3--manifolds bounding 4--manifolds, which draws on the same framework and methods.

Let $\Sigma$ be an oriented surface and $f:\Sigma\to\R$ a Morse function with distinct critical values.
We begin by examining the preimages of points $x\in\R$, denoting the space $f\inv(x)$ by $\Sigma_x$.
Denote a small interval around $x$ by $\varepsilon(x)=(x-\varepsilon,x+\varepsilon)$, and denote the neighbourhood in $\Sigma$ around $\Sigma_x$ by $f\inv(\varepsilon(x))=\varepsilon(\Sigma_x)$.
The spaces $\Sigma_x$ and $\varepsilon(\Sigma_x)$ may not be connected, so we index the connected components by superscript.
Let $p$ be a critical point of $f$ with critical value $f(p)=x$.
By our assumption that critical values are distinct, $p$ is the only critical point of $f$ in $\Sigma_x$.
The connected component of $\Sigma_x$ containing $p$ is called the \emph{critical level} at $p$ and is denoted $\Sigma_x^p$.
The connected component of $\varepsilon(\Sigma_x)$ containing $p$ is called the \emph{critical neighbourhood} at $p$ and is denoted $\varepsilon^p(\Sigma_x)$.
By the regular value theorem, any regular value pulls back through $f\inv$ to a disjoint collection of oriented circles in $\Sigma$.
Likewise, a neighbourhood $\varepsilon(x)$ that contains no critical values of $f$ pulls back to a disjoint collection of annuli in $\Sigma$.
The restriction of $f$ to the components of $\Sigma_x$ that do not contain a critical point has $x$ as a regular value, so the remaining connected components of $\Sigma_x$ are all copies of $S^1$ that we index by $\Sigma_x^i$ for $i=1,\dots,k$, and their associated neighbourhoods are the annuli $\varepsilon^i(\Sigma_x)$.

The shape of a critical level $\Sigma_x^p$ is determined by the index of $p$.
Because $f$ is a Morse function whose domain is a surface, its critical points are easily classified by the Morse lemma (Lemma \ref{lem:morselemma}).
Locally in $\Sigma$, a critical point of index 0 is a minimum, of index 1 a saddle, and of index 2 a maximum of $f$.

Let $x$ be a critical value for the critical point $p$ and take $\varepsilon$ to be small enough that $\varepsilon(x)$ contains no critical values other than $x$.
When $p$ is of index 0 or 2, we can immediately deduce that $\varepsilon^p(\Sigma_x)$ is diffeomorphic to a disc.
When $p$ is of index 1, the Morse lemma tells us that $\Sigma$ looks like a standard saddle near $p$.
The intersection of $\Sigma_x^p$ with this saddle is a cross whose centre is $p$.
For $y\in\varepsilon(x)$, $y$ is a regular value whose preimage is a disjoint union of oriented circles.
The circles above and below the saddle singularity are the result of smoothing out the cross into a pair of oriented arcs, done in two possible ways.
The orientations of these circles orient the cross, which has two incoming arms and two outgoing arms which appear in alternating order.
A Morse function has distinct critical levels, so the cross we know about in $\Sigma_x^p$ must have its arms connected in $\Sigma_x^p$ through nonsingular orientation--preserving arcs.
We can then see that $\Sigma_x^p$ is a figure 8, and $\varepsilon^p(\Sigma_x)$ is a is a pair of pants in $\Sigma$.

This analysis is sufficient to form a handle decomposition for a cobordism of $(\Sigma,\emptyset)$.
We begin with the 3--manifold $\Sigma\times\I$ with boundary components $\Sigma_0=\Sigma\times\{0\}$ and $\Sigma_1=\Sigma\times\{1\}$, and let $f:\Sigma_0\to\I$ be a Morse function with distinct critical values $x_i$.
The general idea is that we attach handles to $\Sigma_0$, altering that boundary component until it is empty.
The $x_i$ partition $\I$ into open intervals containing only regular values.
We take the associated regular annuli to be attaching regions for 3--dimensional 2--handles, and then fill in what remains with 3--dimensional 3--handles.

For an interval $(x_i,x_{i+1})$, consider the subinterval $\varepsilon(t_i)$ where $x_i<t_i<x_{i+1}$ and $\varepsilon$ is small enough that $\varepsilon(t_i)$ contains neither $x_i$ nor $x_{i+1}$.
Take the associated regular annuli $\varepsilon(\Sigma_{t_i})$ to be the attaching regions for 3--dimensional 2--handles.
The boundary component that was once $\Sigma_0$ is now the result of removing the interiors of the annuli $\varepsilon(\Sigma_{t_i})$ from $\Sigma_0$ for each $t_i$, and then introducting discs parallel to the cores of the attached 2--handles.
These discs are seen as $\DD\times\{0\}$ and $\DD\times\{1\}$ inside of a 2--handle $\DD\times\D^1$.
The boundary circles of the discs are $\Sigma_{t_i-\varepsilon}$ and $\Sigma_{t_i+\varepsilon}$ for each $i$.

There are now intervals about each critical point that we know, from our analysis above, pull back to discs, annuli, and pairs of pants.
Each of these subsurfaces have circle boundaries that correspond to the circles $\Sigma_{t_i-\varepsilon}$ and $\Sigma_{t_i+\varepsilon}$ which were capped with discs in the previous step.
The altered boundary described before is therefore a collection of copies of $S^2$, which we take to be the attaching regions of 3--dimensional 3--handles.

With these handles attached, we obtain $(\Sigma\times\I)\cup\{\textrm{2--handles}\}\cup\{\textrm{3--handles}\}$, which is a 3--manifold with exactly one boundary component of $\Sigma_1$, i.e. a cobordism of the pair $(\Sigma,\emptyset)$.

The corresponding dual handle decomposition is realized by turning the process upside down.
Where we previously had 3--handle attachments with cores $\D^3\times\{\vec{0}\}$ and cocores $\{\vec{0}\}\times\{\vec{0}\}$, we now have 0--handles with cores $\{\vec{0}\}\times\{\vec{0}\}$ and cocores $\D^3\times\{\vec{0}\}$.
In other words, the dual construction begins by taking the disjoint union of a collection of 3--discs that correspond to the space obtained by capping the boundary circles of the components $\varepsilon(\Sigma_{x_i})$ with 2--discs and then filling the spheres with 3--balls.
Where we previously had 2--handle attachments with cores $\DD\times\{\vec{0}\}$ and cocores $\{\vec{0}\}\times\D^1$, we now have 1--handle attachments with cores $\{\vec{0}\}\times\D^1$ and cocores $\DD\times\{\vec{0}\}$.
In other words, we connect the 0--handles together using 1--handles.
The old belt 0--spheres of the 2--handles in the previous construction correspond here to new attaching 0--spheres, and the attaching maps are chosen to preserve orientability cf.\ Remark~\ref{rmk:1handle}.
The new attaching 0--spheres bound the new cores, the old core 2--discs are now cocores.
This construction yields a 3--manifold whose boundary is exactly $\Sigma$, and is indeed another cobordism of the pair $(\Sigma,\emptyset)$.



A Stein factorization $f=g\comp h$ is simple to describe.
We first define an equivalence relation on the points in $\Sigma$.
 $p\sim q$ if $p$ and $q$ are 

We first define $h$ on the subspaces $\Sigma_x^i$, i.e.\ the connected levels of $\Sigma$ that do not contain critical points.
Between every pair of adjacent critical values $x_j < x_k$ there is a maximal interval $(x_j,x_k)$ that contains only regular values.
This can be written as $\varepsilon(x)$ for $\varepsilon=(x_k-x_j)/2$ and $x=(x_j+x_k)/2$, and the associated neighbourhood $\varepsilon(\Sigma_x)$ in $\Sigma$ is a collection of annuli that we call \emph{regular annuli}.
Define $h$ to map each component $\Sigma_x^i$ to a distinct value, and require that $h$ be continuous on each regular annulus $\varepsilon^i(\Sigma_x)$.

At this point, $h$ is defined everywhere except for the subspaces $\Sigma_x$ where $x$ is a critical value of $f$, and the Stein complex $h(\Sigma)$ is consists entirely of line segments.
For a critical value $x=f(p)$, the neighbourhood $\varepsilon^p(\Sigma_x)$ of $\Sigma_x$ in $\Sigma$ is a critical neighbourhood along with a collection of annuli.
These connected components intersect with regular annuli on either side of $\Sigma_x$.
A circle $\Sigma_x^i$ is the boundary of exactly two regular annuli $\varepsilon^j(\Sigma_y)$ and $\varepsilon^k(\Sigma_z)$.

Define $h$ on a circle $\Sigma_x^i$ to be a point that connects the line segments associated to the pair of annuli that $\varepsilon^i(\Sigma_x)$ intersects.
Define $h$ on a critical level $\Sigma_x^p$ 

, so $h$ is defined everywhere except for $\Sigma_x$
The circles $\Sigma_x^i$ in $\Sigma_x$ containing no critical points are 


Take $h$ to be a function that crushes the connected components of $\Sigma_x$ to distinct points subject to the constraint that $h(\Sigma_x^i)=h(\Sigma_y^j)$ when $\Sigma_x^i$ and $\Sigma_y^j$ are 



The components of $\varepsilon(\Sigma_x)$ that contain no critical points of $f$ are disjoint unions of annuli, one for each circle in $\Sigma_x$.
Similarly, the preimage of an interval of regular values in $\R$ is the disjoint union of annuli in $\Sigma$.

\{figure: smoothing a cross into a saddle\}

\{figure: neighbourhoods of critical points (index 0,1,2)\}



Let $\Sigma\times I$ be a 3--manifold with oriented boundary components $\Sigma\times\{0\}=\Sigma_0$ and $\Sigma\times\{1\}=\Sigma_1$, where $\Sigma_0$ has orientation opposite that of $\Sigma$ and $\Sigma_1$.
Let $f$ be a Morse function on $\Sigma_0$.
We use a Stein complex of $f$ to build a handle decomposition for a cobordism of the pair $(\Sigma_1,\emptyset)$.
Take $f=g\comp h$, where $g$ is the map that crushes the connected components of $\Sigma_x$ to distinct points, and $h$ is the map that 


 Specify a regular value of $f$ in each connected component of $f(\Sigma_0)$ minus the critical values of $f$.
Each of these regular values has preimage a disjoint collection of circles in $\Sigma_-$ which we take to be the attaching spheres of 3--dimensional 2--handles. 
The resulting object is a 3--manifold with one boundary component of $\Sigma_+$ and the rest corresponding to singular values of $f$.
To obtain a 3--manifold whose boundary is $\Sigma_+$, we need only fill the boundary components corresponding to the singular values of $f$.

\{FIGURE: MORSE FUNCTION AND 2--HANDLE ATTACHMENT\}



We return now to the 3--manifold constructed from $\Sigma\times I$ by gluing 2--handles to the $\Sigma_-$ boundary component.
The boundary components consist of $\Sigma_+$ as well as a collection of components corresponding to the critical points of $f$ left over in the $\Sigma_-$ side of the boundary.
If $p$ has index 0 or 2, then the component of $f\inv(\varepsilon(x))$ contianing $p$ is a disc.
For some $\epsilon$, the boundary of this disc meets the belt sphere of the nearest 2--handle, and the boundary component is seen to be a 2--sphere, which becomes the attaching sphere of a 3--handle.
Similarly, if $p$ has index 1 then the pair of pants near the saddle singularity of $p$ is grown until it meets three belt spheres from three separate 2--handle attachments.
The boundary component is again seen to be a 2--sphere, so we can attach another 3--handle.
It is reasonable to note that, had $\Sigma$ not been oriented, then the boundary component corresponding to a saddle singularity could have been a copy of $\RP^2$.
As $\RP^2$ has Euler characteristic 1, it does not bound any 3--manifold.

\{FIGURE: THREE TYPES OF 3--BALL BLOCKS\}
