\begin{defn}[Smooth Map]
	\label{def:smoothmap}
	Let $(X,\{U_\alpha,\varphi_\alpha\})$ and $(M,\{V_\beta,\psi_\beta\})$ be smooth $n$-- and $k$--manifolds, with $\{U_\alpha,\varphi_\alpha\}$ and $\{V_\beta,\psi_\beta\}$ maximal smooth atlases.
	Let $f:X\to M$ be a map.
	If, for any $\alpha$ and $\beta$, the composition $\psi_\beta\comp f\comp \varphi_\alpha\inv$ is smooth as a map $$\psi_\beta\comp f\comp \varphi_\alpha\inv:\R^n\supset\varphi_\alpha(U_\alpha)\to\varphi_\beta(V_\beta)\subset\R^k,$$ then we say $f$ is \emph{smooth} as a map between manifolds.
	The space of smooth maps $X\to M$ is denoted $C^\infty(X,M)$.
	
	If $f$ is smooth and a well-defined $f\inv$ exists and is smooth, then $f$ is called a \emph{diffeomorphism} between manifolds.
	We say that manifolds are \emph{diffeomorphic} if there exists a diffeomorphism between them.
\end{defn}

\begin{prop}
	\label{prop:diffeoequiv}
	Diffeomorphism is an equivalence relation on the space of smooth manifolds.
\end{prop}

Manifolds are defined by their local homogeneity, and a tangent space is a precise description of that homogeneity near a point.
There are many ways to define tangent spaces, and we do so using derivations as in \cite{Lee00}.
This approach is used for two reasons.
First, it causes the vector space structure of the tangent space to follow directly from the definition.
Second, the definition is applicable to manifolds with faces without any fiddling around with details, and the tangent spaces obtained are still vector spaces of an appropriate dimension.

\begin{defn}[Derivations on $\R^n$]
	Let $x$ be a point in $\R^n$, and denote the space of smooth functions $\R^n\to\R$ be $C^\infty(\R^n)$.
	A linear map $A:C^\infty(\R^n)\to\R$ is called a \emph{derivation at $x$} if it satisfies the product rule
	\[
		A(\varphi\,\psi)=\varphi(x)A\psi+\psi(x)A\varphi
	\]
	for any $\varphi,\psi\in C^\infty(\R^n)$.
	Denote the set of all derivations at $x$ of $C^\infty(\R^n)$ by $T_x \R^n$.
	The derivations $T_x \R^n$ form a vector space under the operations
	\[
		\begin{array}{rcl}
			(A+B)\varphi&=&A\varphi+B\varphi,\\
			(cA)\varphi&=&c(A\varphi).
		\end{array}
	\]
\end{defn}

\begin{prop}[Corollory 3.3 in \cite{Lee00}]
	\label{prop:cor33}
	For any $x$ in $\R^n$, the $n$ derivations
	\[
		\restr{\frac{\pd}{\pd x_1}}{x}, \dots, \restr{\frac{\pd}{\pd x_n}}{x}
	\]
	defined by
	\[
		\restr{\frac{\pd}{\pd x_i}}{x}\varphi=\frac{\pd \varphi}{\pd x_i}(x)
	\]
	form a basis for $T_x\R^n$, and $T_x\R^n$ is therefore an $n$--dimensional vector space. 
\end{prop}

Our definition of derivation extends to the spaces $\HH^k\times\R^{n-k}$, and Proposition~\ref{prop:cor33} extends as well after we establish the following lemma.

\begin{lem}[Extension Lemma, 2.7 in \cite{Lee00}]
	\label{lem:ext}
	Let $M$ be a smooth manifold, let $V\subset M$ be a closed subset, and let $f:V\to\R^d$ be a smooth function.
	Let $U$ be any open set containing $V$.
	Then there exists a smooth extension $\tilde{f}:M\to\R^d$ such that $\restr{\tilde{f}}{V}=f$ and $\supp \tilde{f}\subset U$.
\end{lem}

\begin{defn}[Derivations on $\HH^n$]
	Let $x$ be a point in $\HH^n$.
	A linear map $A:C^\infty(\HH^n)\to\R$ is called a \emph{derivation at $x$} if it satisfies the product rule
	\[
	A(\varphi\,\psi)=\varphi(x)A\psi+\psi(x)A\varphi
	\]
	for any $\varphi,\psi\in C^\infty(\HH^n)$.
	Denote the set of all derivations at $x$ of $C^\infty(\HH^n)$ by $T_x \HH^n$.
	The derivations $T_x \HH^n$ form a vector space under the operations
	\[
	\begin{array}{rcl}
	(A+B)\varphi&=&A\varphi+B\varphi,\\
	(cA)\varphi&=&c(A\varphi).
	\end{array}
	\]
\end{defn}

\begin{prop}[Lemma 3.10 in \cite{Lee00}]
	\label{prop:cor310}
	For any $x$ in $\HH^n$, the $n$ derivations
	\[
	\restr{\frac{\pd}{\pd x_1}}{x}, \dots, \restr{\frac{\pd}{\pd x_n}}{x}
	\]
	defined by
	\[
	\restr{\frac{\pd}{\pd x_i}}{x}\varphi=\frac{\pd \varphi}{\pd x_i}(x)
	\]
	form a basis for $T_x\HH^n$, and $T_x\HH^n$ is therefore an $n$--dimensional vector space. 
\end{prop}

\begin{proof}
	The case where $x$ is in the interior of $\HH^n$ is exactly the same as in Proposition~\ref{prop:cor33}.
	We concern ourselves only with the case when $x$ is a boundary point of $\HH^n$.
	
	Let $i:\HH^n\to\R^n$ be inclusion.
	We will show that $i_*:T_x\HH^n\to T_x\R^n$ is an isomorphism.
	In the case where $i_*A=0$, let $\varphi$ be a smooth, real--valued function defined on a neighbourhood of $x$ in $\HH^n$ and let $\tilde{\varphi}$ extend $\varphi$ to a smooth function on an open subset of $\R^n$.
	From this, we have $\tilde{\varphi}\comp i=\varphi$, so
	\[
		A\varphi = A(\tilde{\varphi}\comp i) = (i_* A)(\tilde{\varphi})=0.
	\]
	This means that $i_*A=0$ implies $A=0$, so $i_*$ is injective.
	For surjectivity, let $B\in T_x\R^n$ and let $\varphi\in C^\infty(\HH^n)$.
	Define $A\in T_x \HH^n$ by
	\[
		A\varphi = B\tilde{\varphi},
	\]
	where $\tilde{\varphi}$ is any extension of $\varphi$.
	Expressing $B$ in terms of the basis for $T_x \R^n$ on the right hand side of this expression gives us
	\[
		A\varphi = \sum_{i=1}^n B_i \frac{\pd \tilde{\varphi}}{\pd x_i}(x).
	\]
	The derivatives $\pd\tilde{\varphi}/\pd x_i$ are determined by the derivatives of $\varphi$ in $\HH^n$, and this definition of $A$ is shown to be a derivation by evaluating $A(\varphi\psi)$ as $B(\tilde{\varphi}\tilde{\psi})$.
	Confirming that $i_*A=B$ uses $\varphi=\tilde{\varphi}\comp i$ as above, and $i_*$ is therefore surjective.
	We conclude that $i_*$ is an isomorphism.
\end{proof}

Note that $\HH^n$ contains model corner points for every type of $j$--corner, $0\leq j\leq n$, along the intersections of its coordinate planes.
Investigating the tangent spaces in $\HH^k\times\R^{n-k}$ allows us to fully realize tangent spaces to manifolds through the use of charts.

\begin{defn}
	Let $M$ be an $n$--manifold and $p$ a point in $M$.
	A linear map $A:C^\infty(M)\to\R$ is a \emph{derivation at p} if it satisfies
	\[
		A(\varphi\,\psi)=\varphi(p)A\psi+\psi(p)A\varphi
	\]
	for any $\varphi,\psi\in C^\infty(M)$.
	Denote the set of all derivations at $p$ of $C^\infty(M)$ by $T_p M$.
	As in the case with $\R^n$, $T_p M$ is a vector space under
	\[
		\begin{array}{rcl}
			(A+B)\varphi&=&A\varphi+B\varphi,\\
			(cA)\varphi&=&c(A\varphi).
		\end{array}
	\]
	We call $T_p M$ the \emph{tangent space to $M$ at $p$} and call the elements of $T_p M$ \emph{tangent vectors}.
\end{defn}

To find a basis, and therefore dimension, of a tangent space we first define how smooth maps act on tangent vectors.

\begin{defn}[Pushforward]
	Let $f:X\to M$ be a smooth map between manifolds.
	For every $p\in X$, we define the map
	\[
		f_*:T_p X\to T_{f(p)} M
	\]
	by
	\[
		(f_*A)(\varphi) = A(\varphi\comp f).
	\]
	If $\varphi\in C^\infty(X)$ then $\varphi\comp f\in C^\infty(M)$, so $X(\varphi\comp f)$ is, at least, defined.
	The operator is clearly linear and it is a derivation at $f(p)$ by
	\[
		\begin{array}{rcl}
			(f_* A)(\varphi\,\psi)&=&A((\varphi\,\psi)\comp f)\\
			&=&X((\varphi\comp f)(\psi\comp f))\\
			&=&(\varphi\comp f(p))A(\psi\comp f) + (\psi\comp f(p))A(\varphi\comp f)\\
			&=&\varphi(f(p))(f_* A)(\psi) + \psi(f(p))(f_* A)(\varphi).
		\end{array}
	\]
	The map $f_*$ is called the \emph{pushforward} of $f$, and it relates how the tangent vectors in $T_p X$ are transformed into tangent vectors of $T_{f(p)}M$ when $p$ is passed through $f$.
\end{defn}

\begin{lem}[Lemma 3.5 in \cite{Lee00}]
	\label{lem:pushprop}
	Let $f:X\to M$ and $g:M\to N$ be smooth maps.  Let $p\in X$.
	\begin{enumerate}
		\item $f_*:T_p X\to T_{f(p)}M$ is linear.
		\item $(g\comp f)_*=g_*\comp f_*:T_p X\to T_{g\comp f(p)}N$
		\item $(\id_X)_*=\id_{T_p X}:T_p X\to T_p X$
		\item If $f$ is a diffeomorphism then $f_*:T_p X\to T_{f(p)}M$ is an isomorphism.
	\end{enumerate}
\end{lem}

\begin{prop}[Proposition 3.7 in \cite{Lee00}]
	\label{prop:incl}
	Let $M$ be a smooth manifold and $U\subset M$ an open submanifold.
	Let $i:U\to M$ be inclusion.
	For any $p\in U$, the pushforward $i_*:T_p U\to T_p M$ is an isomorphism.	
\end{prop}

We know that $T_x \HH^n$ is $n$--dimensional for any $x$ in $\HH^n$.
Combining Lemma~\ref{lem:pushprop} and Proposition~\ref{prop:incl} with a smooth chart proves the corresponding dimension theorem for tangent spaces to smooth manifolds.

\begin{lem}
	Let $M$ be a smooth $n$--manifold and $p$ a point in $M$.
	Then $T_p M$ is an $n$--dimensional vector space with basis given by the coordinate vectors
	\[
		\restr{\frac{\pd}{\pd x_1}}{p}, \dots,	\restr{\frac{\pd}{\pd x_n}}{p}
	\]	
	in any smooth chart.
\end{lem}

\begin{proof}
	Let $(U,\varphi)$ be a chart about $p$.
	Then $\varphi$ is a diffeomorphism from $U$ to an open subset $\tilde{U}$ of $\HH^k\times\R^{n-k}$ for some value of $k$. 
	By Lemma~\ref{lem:pushprop} and Proposition~\ref{prop:incl}, the pushforward $\varphi_*$ is an isomorphism $T_p M\to T_{\varphi(p)}\HH^k\times\R^{n-k}$.
	By Proposition~\ref{prop:cor310} $T_x\HH^k\times\R^{n-k}$ is an $n$--dimensional vector space with basis the derivations
	\[
		\restr{\frac{\pd}{\pd x_1}}{x}, \dots, \restr{\frac{\pd}{\pd x_n}}{x},
	\]
	so $T_p M$ is as well.
\end{proof}

We also classify smooth maps between manifolds.
This classification helps us define a submanifold and provides some of the groundwork for defining handles in the next section.

\begin{defn}[Embedding]
	Let $f:X\to M$ be a smooth map between manifolds.
	The \emph{rank} of $f$ at the point $p\in X$ is the rank of the pushforward $f_*$ as a linear map.
	This is computed as the rank of the matrix of partial derivatives of $f$ in any smooth chart or as the dimension of $f_*(T_p X)\subset T_{f(p)} M$.
	If $f$ has rank $k$ for every $p$ in $X$, then $f$ is of \emph{constant rank} and we say $\rk f=k$.
		
	If $f_*$ is injective at $p$, then $f$ is \emph{immersive} at $p$.
	If $f$ is an everywhere immersive, then it is an \emph{immersion}.
	This is equivalent to $\rk{f}=\dimn X$.
	If $f_*$ is surjective at $p$, then $f$ is \emph{submersive} at $p$, and $p$ is a \emph{regular point} of $f$.
	If $f$ is everywhere submersive, then it is a \emph{submersion}.
	This is equivalent to $\rk{f}=\dimn M$.
	If $\rk{f_*}$ is less than maximal, then $p$ is a \emph{critical point} of $f$.
	In this case, $f(p)=q$ is a \emph{critical value}.
	The set of critical points of $f$ in $M$ is called the \emph{singular set} of $M$.
	If $p$ is a critical point and the Hessian of $f$, i.e.\ the matrix of second partial derivatives of $f$ in any smooth chart, is of less than full rank, then $p$ is a \emph{degenerate critical point}.
	For $q\in M$, the set of points $x$ in $X$ such that $f(x)=q$ is called the \emph{preimage} of $q$ through $f$ or the \emph{fiber} of $f$ over $q$ and is denoted $f\inv(q)$.
	A point $q \in M$ whose fiber consist entirely of regular points is a \emph{regular value} of $f$.
	
	If $f$ is an injective immersion that is a homeomorphism onto its image $f(X)\subset M$ under the subspace topology, then it is a \emph{smooth embedding} or just \emph{embedding}.
	
	Let $X\subset M$ where $M$ is a manifold, and let $i:X\into M$ be inclusion.
	If $i$ is an embedding, then we call $X$ an \emph{embedded submanifold} or \emph{submanifold} of $M$, and $X$ is, by itself, a manifold with smooth structure induced by the embedding.
\end{defn}

\begin{prop}
	\label{prop:boundariesaremanifolds}
	Let $M$ be an $(n+1)$--manifold.
	Any connected face in $\pd_f M$ is an $n$--dimensional submanifold of $M$.
\end{prop}

\begin{theorem}[Regular Value Theorem]
	Let $f:X\to M$ be a smooth map between manifolds.
	Let $q\in M$ be a regular value.
	The preimage $f\inv(q)$ is a $(\dim X-\dim M)$--submanifold of $X$.
\end{theorem}

Maps that are ``close enough'' to one another via perturbation tend to be considered equivalent.
This concept is made precise through homotopy.

\begin{defn}[Homotopy]
	\label{def:homotopy}
	Let $f,g:X\to M$ be smooth maps between smooth manifolds.
	Denote the closed unit interval $[0,1]$ by $\I$.
	Let $H$ be a continuous function 
	\[
		\begin{array}{cccc}
			H: & X\times \I & \to & M \\
			& (x,t)	& \mapsto & H_t(x)
		\end{array}
	\]
	with $H_t$ a continuous function $X\to M$ for all $t$, $H_0(x)=f(x)$, and $H_1(x)=g(x)$.
	Then $H$ is called a \emph{homotopy} between $f$ and $g$.
	If a homotopy exists, then $f$ and $g$ are \emph{homotopic}.
	Less formally, $f$ and $g$ being homotopic means that one can be continuously deformed into the other.
	The topological spaces $X$ and $M$ are \emph{homotopy equivalent} if there exist continuous maps $f:X\to M$ and $g:M\to X$ for which $g\comp f$ is homotopic to $\ident{X}$ and $f\comp g$ is homotopic to $\ident{M}$.
	
	Because we are primarily interested in smooth functions to build our machinery, we extend our definition of homotopy to a smooth version.
	With the notation above, a smooth map $H:X\times\I\to M$ with $H_0(x)=f(x)$ and $H_1(x)=g(x)$ is a \emph{smooth homotopy} between $f$ and $g$.
	If a smooth homotopy exists, then $f$ and $g$ are \emph{smoothly homotopic}.
	
	Let $M$ be a manifold and $X\subset M$ a subspace that is not necessarily a submanifold.
	A continuous map $F:M\times\I\to M$ such that, for every $m$ in $M$ and $x$ in $X$, $F(m,0)=m$, $F(m,1)\in X$, and $F(x,1)=x$, is called a \emph{deformation retraction}, and the subspace $X$ is called a \emph{deformation retract} of $M$.
	Because a deformation retraction is a homotopy equivalence between $M$ and $X$, a deformation retraction from $M$ onto $X$ exists if and only if $M$ and $X$ are homotopy equivalent.
\end{defn}

\begin{prop}
	For a fixed pair $(X,M)$ of spaces, homotopy and smooth homotopy are equivalence relations on the space of maps $X\to M$.
\end{prop}

\begin{defn}
	Let $X$ be a topological space, let $P$ be a property that an element of $X$ can possess, and let $X_P$ be the subset of elements of $X$ that posses property $P$.
	A property $P$ is \emph{stable under perturbation} if, for any element $x$ in $X_P$, there is an open set $U$ of $X$ such that $x\in U$ and $U\subset X_P$.
	A property is \emph{generic} if $X_P$ is open and dense in $X$
\end{defn}

We are mainly concerned with the property of having no degenerate critical points.
Consider the map $f:\R\to\R$, $f(x)=x^3$.
The critical point at $x=0$ is degenerate, but the maps $H_{\varepsilon}(x)=x^3+\varepsilon x$ have either a pair of nondegenerate critical points (for $\varepsilon<0$) or no critical points (for $\varepsilon>0$), and in either case $H_{\varepsilon}(x)$ has no degenerate critical points and that property is stable.
Degenerate critical points are easily ``fixed,'' so our model generic functions are those without degenerate critical points.

\begin{prop}
	A generic smooth map between manifolds has no degenerate critical points.
\end{prop}

Intersections between embedded submanifolds also fit in the language of genericity.
We call this property transversality.

\begin{defn}
	Let $M$ be an $n$--manifold with submanifolds $X$ and $Y$.
	We say that $X$ and $Y$ \emph{intersect transversely} as submanifolds if, at every intersection point, the sum of their tangent spaces is the tangent space of $M$ at that point.
	Symbolically, transversality is denoted by $\pitchfork$ and the condition is stated by $X\pitchfork Y$ if and only if for every point $p\in X\cap Y$, $T_p X+ T_p Y=T_p M$.
	We may also say that an intersection point is, itself, \emph{transverse}, or that $X$ and $Y$ \emph{intersect transversally at a point}.	
\end{defn}

Transversality of a pair of submanifolds $X$, $Y$ is stable under small perturbations of the embeddings $X\into M$ and $Y\into M$.
A point of $X\cap Y$ that is not a transverse intersection may be perturbed into stable transversality or non-intersection, so we say that intersections of submanifolds with complementary codimension are generically transverse.

\begin{prop}
	If $X$, $Y$ are generic embedded submanifolds of $M$, then they intersect transversely.
\end{prop}

The main result of this work is applicable to orientable 3--manifolds, so we will review what is meant by a space being orientable.
Orientability is the guarantee that when you go for a walk through a manifold you don't return home to find that your right and left sides have switched places.

We first define orientations on real vector spaces.
Orientations of real vector spaces let us define orientations of manifolds.
Orientations of manifolds let us define a smooth map as orientation preserving or orientation reversing.

\begin{defn}
	\label{def:orientation}
	Let $V^n$ be an $n$--dimensional real vector space and $\mathcal{B}(V^n)$ the set of ordered bases for $V^n$.
	Then $\mathcal{B}(V^n)$ has exactly two path--connected components.
	An \emph{orientation} of $V^n$ is a choice of one of these components.
	
	Let $A:V\to W$ be a linear map between $n$--dimensional oriented real vector spaces.
	If $A$ maps the orientation of $V$ to the orientation of $W$, $A$ is \emph{orientation preserving}.
	Otherwise, $A$ maps the the orientation of $V$ to the orientation opposite that of $W$, and in this case $A$ is \emph{orientation reversing}.	
\end{defn}

\begin{ex}
	Consider $V=\HH^k\times\R^{n-k}$.
	At $x\in V$, the basis vectors of $T_x V$ are the derivations
	\[
		\restr{\frac{\pd}{\pd x_1}}{x}, \dots, \restr{\frac{\pd}{\pd x_n}}{x}.
	\]
	This means that there is a canonical orientation of $T_x V$ induced by the orientation of $V$ for every $x\in V$.
	Now, let $y\in V$ different from $x$.
	Because all of the tangent spaces of $V$ have the same basis, an orientation of $T_x V$ can be meaningfully compared to an orientation of $T_y V$.
	In particular, we can determine whether $T_x V$ and $T_y V$ have the same orientation.
\end{ex}

The fact that orientations of tangent spaces in $\HH^k\times\R^{n-k}$ can be compared leads to our definition of orientation for manifolds.
We define an orientation of a manifold $M$ to be a continuous orientation of the tangent spaces to $M$.
We use continuous here to mean that the orientation is constant on any chart.

\begin{defn}
	Let $M$ be an $n$--manifold with maximal smooth atlas $\mathcal{A}$, let $(\varphi, U)\in\mathcal{A}$.
	A \emph{continuous choice of orientation} on $U$ is an orientation of every tangent space to $U$ such that, for every $p,q\in U$, the linear isomorphisms
	\begin{eqnarray*}
		\varphi_*^p:T_p U&\to& T_{\varphi(p)}\varphi(U)\\
		\varphi_*^q:T_q U&\to& T_{\varphi(q)}\varphi(U)		
	\end{eqnarray*} take the orientations of $T_p U$, $T_q U$ to the same orientation of $T_{\varphi(p)}\varphi(U)\cong  T_{\varphi(q)}\varphi(U)$.
	An \emph{orientation} of $M$ is an orientation of every tangent space to $M$ that induces a continuous choice of orientation for every chart of $\mathcal{A}$. 
\end{defn}

\begin{defn}
	Let $f:X\to M$ be a smooth map between oriented manifolds.
	If $f_*$ is orientation preserving (reversing) for every $p\in X$, $f$ is an \emph{orientation preserving (reversing)} smooth map.	
\end{defn}