The \emph{CW--complex} is our first foray into combinatorial descriptions of manifolds.

\begin{defn}
	For $n>0$, a topological space that is homeomorphic to the $n$--ball is called an \emph{open $n$--cell}, and a topological space that is homeomorphic to the $n$--disc is called a \emph{closed $n$--cell}.
	For $n=0$, we say that a singleton set is both a closed and open 0--cell.
	
	Let $M$ be a Hausdorff space.
	A \emph{cell decomposition} of $M$ is a collection $T=\{t_\alpha\}$ such that every $t_\alpha$ is an open cell, every pair of cells of $T$ is disjoint, and the union of the cells of $T$ is exactly $M$.
	The \emph{$n$--skeleton} of $T$ is the disjoint union of the cells of $T$ with dimension at most $n$, and it is denoted $T^n$.
	
	Let $T$ be a cell decomposition of the Hausdorff space $M$.
	Then $T$ is a \emph{closure--finite weak--topology complex} if the following are satisfied: 
	\begin{enumerate}
		\item For any $n$--cell $t_\alpha$ in $T$, there exists a continuous map $f:\DN\to M$ such that $\restr{f}{B^n}$ is a homeomorphism onto $t_\alpha$ and $f(S^{n-1})$ is contained in $T^{n-1}$.
		\item The closure of any cell of $T$ intersects only a finite number of other cells in $T$ (i.e.\ $T$ is closure finite).
		\item A subset $X$ of $M$ is closed if and only if $X\cap \overline{t_\alpha}$ is closed for every $t_\alpha$ in $T$ (i.e. the topology of $M$ is weak).		
	\end{enumerate}
	It is standard to abbreviate such a cell decomposition as a \emph{CW--complex}.	
\end{defn}

We concern ourselves mostly with a specific type of cell decomposition called a \emph{triangulation.}
The cells of a triangulation are the interiors of \emph{$n$--simplices}, which can be through of as generalizations of the most basic 2--dimensional building block, the triangle, to arbitrary dimensions.

\begin{defn}
	Let $E=\{e_0,\dots,e_n\}$ be a set of $n+1$ points in some $n$--dimensional affine space $F$ such that
	\[
	  e_i\neq \sum_{j\neq i}t_j e_j
	\]
	for each $i$ and any collection of nonnegative real $t_j$ with $\sum t_j=1$.
	The \emph{standard} $n$--\emph{simplex} $\sigma$ is defined as
	\[
	  \sigma = \{p\in F : p = \sum_{i=1}^n t_i e_i,\text{ } e_i\in E,\text{ } t_i\geq 0\text{ for each }i, \text{ and }\sum_{i=1}^n t_i=1\}.
	\]
	The above construction is called taking the \emph{convex hull} over $E$.
	
	Let $E'$ be a $k+1$ element subset of $E$.
	The convex hull over $E'$ is called a \emph{facet} or \emph{$k$--facet} of $\sigma$ and is itself a $k$--simplex.
	A zero--dimensional facet is a \emph{vertex}, one--dimensional an \emph{edge}, two--dimensional a \emph{triangle}, three--dimensional a \emph{tetrahedron} and four--dimensional a \emph{pentachoron}.
	These names are also applied to the 0--, 1--, 2--, 3--, and 4--simplices.
	A facet with dimension $n-1$ is called a \emph{face} of the simplex that contains it.
	We number the vertices of the $n$--simplex with the numbers $0,\dots,n$.
	Every face of $\sigma$ contains all but one vertex of $\sigma$, and this gives a numbering to the faces of $\sigma$.
	The $i^{th}$ face of $\sigma$ is found via the \emph{face map} $F^i(\sigma)$.
	Removing all of the proper faces of a simplex leaves us with the \emph{interior} of $\sigma$.
\end{defn}

Let's start building some cell structures out of simplices.
First, notice that a simplex has a canonical cell decomposition into the interiors of its facets.
Next, let $\sigma$, $\tau$ be $n$--simplices, $n>0$.
A linear attaching map $F^i(\sigma)\to F^j(\tau)$ can be defined by mapping the vertices of $F^i(\sigma)$ to the vertices of $F^j(\tau)$ and then extending linearly over the facets of $F^i(\sigma)$.
Let $g:F^i(\sigma)\to F^j(\tau)$ be such an attaching map, and form the space $\sigma\cup_g \tau$.
There is also a canonical cell decomposition of $\sigma\cup_g \tau$ built from the interiors of the facets of $\sigma$ and $\tau$.
Extending to a collection of $n$--simplices and gluing maps, a cell decomposition continues to be simple to define, and we may say the the adjunction formed from simplices in a collection $\{\sigma_i\}$ over gluing maps $\{g_j\}$ has a canonical cell decomposition.

\begin{defn}
  A CW--complex matching the description above is called an \emph{$n$--dimensional simplicial gluing complex} or just \emph{gluing complex}.
  The faces of a gluing complex $T$ can be partitioned into two subsets based on whether the face has been glued to another face.
  If a face $F^i(\sigma)$ has been glued to another face $F^j(\tau)$, both faces are \emph{glued}.
  Otherwise, $F^i(\sigma)$ is \emph{unglued}.
  The set of unglued faces and all facets contained in those faces form the \emph{boundary} of a gluing complex, and is denoted $\pd T$.
  The remaining facets are \emph{internal}.
  A gluing complex with empty boundary is \emph{closed}.

  Put $n=k+m$ and let $\sigma$ be an $m$--simplex in an $n$--gluing complex $T$ and consider an $n$--simplex $\tau$ containing $\sigma$.
  There is a $k$--facet $\tau_k$ that is ``opposite'' $\sigma$ in the sense that $\sigma\cap\tau_k=\emptyset$ and $\tau$ is the convex hull of $\sigma\cup\tau_k$.
  The subset of $C$ built as the union of all $\tau_k$ corresponding to $\tau$ containing $\sigma$ is the \emph{link} of $\sigma$ and is denoted $\lk{\sigma}$.
  Note that $\lk{\sigma}$ is an $m$--gluing complex in its own right.

  If $M$ is an $n$--manifold and $T$ is an $n$--simplicial gluing complex that is homeomorphic to $M$, then we say that $T$ is a \emph{triangulation} of $M$.
\end{defn}

There is a very slick way of checking whether a gluing complex is a triangulation of some $n$--manifold using only the links of vertices.

\begin{theorem}
  An $n$--gluing complex $T$ is the triangulation of some $n$--manifold $M$ if and only if the link of every internal vertex $v$ is homeomorphic to $S^{n-1}$ and the link of every boundary vertex $w$ is homeomorphic to $D^{n-1}$ with $\pd\lk{w}$ contained in $\pd T$.
\end{theorem}

When we have a triangulation $T$ for a closed, oriented $n$--manifold $M$, we can also define a structure dual to that triangulation that is useful for computation.
This structure, denoted $T^*$, is a cell decomposition dual to $T$ in the sense that each $(n-k)$--cell of $T^*$ is associated with a $k$--cell of $T$.
For a $k$--simplex $\sigma$ of $T$, denote the dual cell in $T^*$ to $\sigma$ by $\sigma^*$.
When $T^*$ is fully defined, this will form a bijection between the $k$--cells of $T$ and the $(n-k)$--cells of $T^*$.
Now, $\sigma$ is contained in a finite number of $n$--simplices $\tau_i$, and we can fully describe $\sigma^*$ by describing how $\sigma^*$ meets these $\tau_i$.

For a given $\tau$ containing $\sigma$, there are $2^{n-k}$ subsets of the vertices of $\tau$, i.e.\ the 0--skeleton of $\tau$, i.e.\ $\tau^0$, that contain the vertices of $\sigma$.
Each subset $s_j\subset\tau^0$ has a barycentre $b(s_j)$, so we form a collection $B_\tau(\sigma)=\{b(s_j)\}$ of these barycentres.
Then the intersection of $\sigma^*$ with $\tau$ is defined as the convex hull over $B_\tau(\sigma)$ inside of $\tau$.
The union of these hulls over all $\tau_i$ containing $\sigma$ is then a closed $(n-k)$--cell whose interior is then defined to be the dual cell $\sigma^*$.
The union of all $\sigma^*$ forms a CW--complex $T^*$ for $M$ that we call the \emph{dual decomposition} to $T$.








%A theorem of Whitehead \{CITE: Whitehead 1940\} tells us that every smooth manifold has a canonical piecewise--linear structure hence a structure as a simplicial manifold.
%It is also true that every simplicial manifold of dimension $\leq 6$ has an essentially unique smoothing up to diffeomorphism \{CITE:Milnor, topology 46 years later\}
